---
title: "Regularization"
author: "Kendra Wyant"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: 
  html_document:
    toc: true 
    toc_depth: 4
---

### Notes
Purpose: Demo of different regularization techniques using tidymodels in the context 
of a full workflow.   


### Setup
```{css, include = FALSE}
pre, code {
  max-height: 500px;
  overflow-y: auto;
  white-space: pre !important; 
  overflow-x: auto
}
```

Paths 
```{r}
path_data <- ""
```

Packages and Source
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(tidymodels)
library(kableExtra)
library(skimr)
library(naniar)
library(doParallel) # for parallel processing
library(mlbench) # where our dataset comes
library(vip)
library(Matrix)
library(glmnet)

options(tibble.print_max = Inf)
```

Setup Parallel Processing
```{r}
n_core <- detectCores(logical = FALSE)

cl <- makePSOCKcluster(n_core - 1)
registerDoParallel(cl)
```

Define function for plotting hyperparameters (Written by John Curtin)
```{r}
plot_hyperparameters <- function(tune_fit, hp1, hp2 = NULL, metric = NULL, log_hp1 = FALSE) {

  data <- collect_metrics(tune_fit)
  
  metric_scores <- data %>% 
    filter(.metric == metric) %>% 
    pull(mean)
    
  x1 <- data[[hp1]]
  if (log_hp1) x1 <- log(x1)
  
  if (is.null(hp2)) {
    ggplot(mapping = aes(x = x1, y = metric_scores)) +
      geom_line() +
      xlab(hp1) +
      ylab(metric)
  } else {
    x2 <- factor(data[[hp2]], ordered = TRUE)
    ggplot(mapping = aes(x = x1, y = metric_scores, group = x2, color = x2)) +
      geom_line() +
      xlab(hp1) +
      ylab(metric) +
      scale_color_discrete(name = hp2)
  }
}
```



### Load Data

```{r}
data("PimaIndiansDiabetes2", package = "mlbench")

data_all <- PimaIndiansDiabetes2 %>% 
  mutate(pregnant_history = ifelse(pregnant == 0, "Never been pregnant", "Has been pregnant"),
         insulin_rating = case_when(insulin < 100 ~ "Average", 
                                    insulin >= 100 & insulin <= 125 ~ "Moderately high",
                                    insulin > 125 ~ "Very high"),
         weight = case_when(mass < 18.5 ~ "Underweight",
                            mass >= 18.5 & mass < 25 ~ "Average",
                            mass >= 25 & mass < 30 ~ "Overweight",
                            mass >= 30 ~ "Obese")) %>% 
  rename(n_pregnancies = pregnant,
         bmi = mass)
```


### Cleaning EDA (looks for errors that need to be cleaned)

**Data leakage** = "When information from outside the training dataset is used to create the model. This additional information can allow the model to learn or know something that it otherwise would not know and in turn invalidate the estimated performance of the mode being constructed."     

Glimpse
```{r}
data_all %>% 
  glimpse()
```

Missing data
```{r}
data_all %>% 
  miss_var_summary()
```

Numeric Predictors
```{r}
data_all %>% 
  select(where(is.numeric)) %>% 
  psych::describe() %>% 
  select(n, min, max)
```

Categorical Predictors
```{r}
data_all %>% 
  select(where(is.character) | where(is.factor)) %>% 
  map(~ distinct(as_tibble(.))) 
```


### Test/train split (also may have validation set)
```{r}
set.seed(20210606)
splits <- data_all %>% 
  initial_split(prop = 3/4, strata = "diabetes")

data_trn <- analysis(splits) %>% 
  glimpse()

data_test <- assessment(splits) %>% 
  glimpse()
```


### Modeling EDA (To inform feature engineering)

Distributions
```{r}
data_trn %>% 
  skim()
```


<br>


Other EDA:    

* Univariate (e.g., bar plots, histograms)   
* Bivariate (e.g., relationship between predictors and outcome, correlation matrix)   


### Our Model 

Outcome = dichotomous    
Statistical algorithm = logistic regression   
Performance metric = accuracy    

Tidymodels workflow:   
1. Create recipe  
2. Make features      
3. Model selection (in training data)   
4. Model evaluation (in test data)   


### Recipe

*Any feature engineering will be done in your recipe using step functions*   

Many different step functions for preprocessing data - https://recipes.tidymodels.org/reference/index.html     

Minimum recipe needed to perform logistic regression
```{r}
rec_lr <- recipe(diabetes ~ ., data = data_trn) %>% 
  step_meanimpute(all_numeric()) %>% 
  step_string2factor(diabetes, levels = c("neg", "pos")) %>%   # set outcome variable to factor
  step_string2factor(all_nominal()) %>%
  step_modeimpute(all_nominal()) %>% 
  step_dummy(all_nominal(), -diabetes) 
```

However, for regularization we want to apply a penalty consistently to all predictors.
So, we need to put all predictors on the same scale using step_normalize
```{r}
rec <- recipe(diabetes ~ ., data = data_trn) %>% 
  step_meanimpute(all_numeric()) %>% 
  step_string2factor(diabetes, levels = c("neg", "pos")) %>%  # set outcome variable to factor
  step_string2factor(all_nominal()) %>%
  step_modeimpute(all_nominal()) %>% 
  step_dummy(all_nominal(), -diabetes) %>% 
  step_normalize(all_predictors())
```



Other steps that might come in handy:   

* step_interact = add interactions to model
* step_YeoJohnson = Yeo Johnson transformation (similar to Box Cox transformation)   
* step_pca = principle components analysis on group of variables   
* step_nzv = removes variables with near zero variance   
* step_mutate = similar to mutate in dplyr and gives you flexibility for creating 
specific features    

<br>

**Question: Why wouldn't we do data processing on the full data set?**      

<br>

### Make features

Training feature matrix   
Features without scaling
```{r}
feat_trn <- rec_lr %>% 
  prep(training = data_trn, strings_as_factors = FALSE) %>% # calculates any stats from recipe
  bake(new_data = data_trn) %>% # creates features in training or test data
  glimpse()
```

Features with scaling    
```{r}
feat_trn <- rec %>% 
  prep(training = data_trn, strings_as_factors = FALSE) %>% # calculates any stats from recipe
  bake(new_data = data_trn) %>% # creates features in training or test data
  glimpse()
```

<br>

*In the real world this process is iterative*  
*Here you will want to do some additional EDA here and update your recipe as needed*  

<br>

Check if outcome is balanced   
```{r}
feat_trn %>% 
  ggplot(aes(x = diabetes)) +
  geom_bar(fill = "light grey", color = "black") +
  theme_classic()
```



Once you have your final recipe you will also want to make a feature matrix from test 
set. **Do not perfom any EDA on this feature matrix though!** This feature matrix ideally 
should be used only once for evaluating your model.    

```{r}
feat_test <- rec %>% 
  prep(training = data_trn, strings_as_factors = FALSE) %>% 
  bake(new_data = data_test) # bake on test data
```





### Model Selection 

Set up a tracking tibble to keep track of model performance
```{r}
track_accuracy <- tibble(model = character(),
                         accuracy_trn = numeric(),
                         n_features = numeric())
```


#### 1. No regularization

Fit model
```{r}
fit_lr <-
  logistic_reg() %>% 
  set_engine("glm") %>% 
  fit(diabetes ~ ., data = feat_trn)
```

estimated model performance
```{r}
lr_train <- accuracy_vec(feat_trn$diabetes, predict(fit_lr, feat_trn, type = "class")$.pred_class)
```

n features
```{r}
lr_feat <- fit_lr %>% 
  tidy() %>% 
  filter(estimate != 0 & term != "(Intercept)") %>% 
  nrow()

# note: to see model coefficients use this code
fit_lr %>% 
  tidy()
```

add to tracking tibble
```{r}
track_accuracy <- add_row(track_accuracy,
                          model = "LR w/no regularization",
                          accuracy_trn = lr_train,
                          n_features = lr_feat)

track_accuracy
```


#### 2. Lasso regularization

Hyperparameter = An L1 norm penalty (λ)  

With regularization we will need to use resampling to tune our hyperparameter
```{r}
splits_boot <- data_trn %>% 
  bootstraps(times = 100, strata = "diabetes")
```

Set tune grid
```{r}
grid_penalty <- expand_grid(penalty = exp(seq(-7, 3, length.out = 200)))
```

Fit models
```{r}
fits_lasso <- logistic_reg(penalty = tune(), 
                           mixture = 1) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  tune_grid(preprocessor = rec,
            resamples = splits_boot,
            grid = grid_penalty,
            metrics = metric_set(accuracy))
```

Check hyperparameters
```{r}
plot_hyperparameters(fits_lasso, hp1 = "penalty", metric = "accuracy", log_hp1 = TRUE)
```

Best penalty value
```{r}
show_best(fits_lasso)
```

estimated model performance   
**Question: Why do we first need to train the model again before getting performance and coefficient estimates?**
```{r}
fit_lasso <- logistic_reg(penalty = select_best(fits_lasso)$penalty,
                        mixture = 1) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  fit(diabetes ~ ., data = feat_trn)


lasso_train <- accuracy_vec(feat_trn$diabetes, predict(fit_lasso, feat_trn, type = "class")$.pred_class)
```

n features
```{r}
# Coefficients have now been zeroed out of the model
fit_lasso %>% 
  tidy() 

lasso_feat <- fit_lasso %>% 
  tidy() %>% 
  filter(estimate != 0 & term != "(Intercept)") %>% 
  nrow()
```

add to tracking tibble
```{r}
track_accuracy <- add_row(track_accuracy,
                          model = "Lasso regularization",
                          accuracy_trn = lasso_train,
                          n_features = lasso_feat)

track_accuracy
```

<br>

Lets see how these 2 models compare in new data  
*Note: this is only for educational purposes, you would not want to do anything with your test set yet*
```{r}
lr_test <- accuracy_vec(feat_test$diabetes, predict(fit_lr, feat_test, type = "class")$.pred_class)
lasso_test <- accuracy_vec(feat_test$diabetes, predict(fit_lasso, feat_test, type = "class")$.pred_class)

track_accuracy %>% 
  mutate(accuracy_test = c(lr_test, lasso_test)) %>% 
  select(1:2, 4, 3)
```

**Question: What might explain these differences in performance in new data?**


#### 3. Ridge regularization

Hyperparameter = An L2 norm penalty (λ)   

We can use same bootstrap splits for our ridge and elasticnet models   


Set tune grid
```{r}
grid_penalty <- expand_grid(penalty = exp(seq(-6, 4, length.out = 200)))
```

Fit models
```{r}
fits_ridge <- logistic_reg(penalty = tune(), 
                           mixture = 0) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  tune_grid(preprocessor = rec,
            resamples = splits_boot,
            grid = grid_penalty,
            metrics = metric_set(accuracy))
```

Check hyperparameters
```{r}
plot_hyperparameters(fits_ridge, hp1 = "penalty", metric = "accuracy", log_hp1 = TRUE)
```

Best penalty value
```{r}
show_best(fits_ridge)
```

estimated model performance
```{r}
fit_ridge <- logistic_reg(penalty = select_best(fits_ridge)$penalty,
                        mixture = 0) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  fit(diabetes ~ ., data = feat_trn)


ridge_train <- accuracy_vec(feat_trn$diabetes, predict(fit_ridge, feat_trn, type = "class")$.pred_class)
```

n features
```{r}
ridge_feat <- fit_ridge %>% 
  tidy() %>% 
  filter(estimate != 0 & term != "(Intercept)") %>% 
  nrow()
```

add to tracking tibble
```{r}
track_accuracy <- add_row(track_accuracy,
                          model = "Ridge regularization",
                          accuracy_trn = ridge_train,
                          n_features = ridge_feat)

track_accuracy
```

<br>

#### 4. Elasticnet regularization 

This model has two hyper-parameters:  

λ controls the degree of regularization as before
α is a “mixing” parameter that blends the degree of L1 and L2 contributions to the 
aggregate penalty. (Proportion of LASSO penalty)  
α = 1 results in the LASSO model   
α = 0 results in the Ridge model   
Intermediate values for α blend these penalties together proportionally to include 
more or less LASSO penalty   


Set tune grid
```{r}
grid_penalty <- expand_grid(penalty = exp(seq(-6, 3, length.out = 200)),
                            mixture = seq(0, 1, length.out = 11))
```

Fit models
```{r}
fits_en <- logistic_reg(penalty = tune(), 
                        mixture = tune()) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  tune_grid(preprocessor = rec,
            resamples = splits_boot,
            grid = grid_penalty,
            metrics = metric_set(accuracy))
```

Check hyperparameters
```{r}
plot_hyperparameters(fits_en, hp1 = "penalty", hp2 = "mixture", metric = "accuracy", log_hp1 = TRUE)
```

Best penalty value
```{r}
show_best(fits_en)
```

estimated model performance
```{r}
fit_en <- logistic_reg(penalty = select_best(fits_en)$penalty,
                       mixture = 0) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  fit(diabetes ~ ., data = feat_trn)


en_train <- accuracy_vec(feat_trn$diabetes, predict(fit_en, feat_trn, type = "class")$.pred_class)
```

n features
```{r}
en_feat <- fit_en %>% 
  tidy() %>% 
  filter(estimate != 0 & term != "(Intercept)") %>% 
  nrow()
```

add to tracking tibble
```{r}
track_accuracy <- add_row(track_accuracy,
                          model = "Elasticnet regularization",
                          accuracy_trn = en_train,
                          n_features = en_feat)

track_accuracy
```

<br>

**Question: Which model should we select?**   

**Why do you think this model performed best?**

<br>

### Model evaluation

Note we are not selecting and evaluating our model on the same data to avoid optimization 
bias. Optimization bias results in slightly overestimating how well a model will perform 
on new data. This may be okay depending on your research question. 

```{r}
fit_best <- logistic_reg(penalty = select_best(fits_en)$penalty,
                       mixture = 0) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  fit(diabetes ~ ., data = feat_trn)

accuracy_vec(feat_test$diabetes, predict(fit_best, feat_test)$.pred_class)
```

Get final model coefficients on full data set
```{r}
# feature matrix
feat_all <- rec %>% 
  prep(training = data_all, strings_as_factors = FALSE) %>% 
  bake(new_data = data_all) 

# refit best model on entire data set
fit_best_full <- logistic_reg(penalty = select_best(fits_en)$penalty,
                       mixture = 0) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") %>% 
  fit(diabetes ~ ., data = feat_all)

fit_best_full %>% 
  tidy()
```

Feature important indices
```{r}
fit_best_full$fit %>% 
  vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
```




